---
title: "Midterm"
author: "Isabella Chittumuri"
date: "10/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2
Textbook pg 7 at last paragrpah 

# 4
you're given an interval and you have a 95% CI that you're 


Maybe
# 4
# 6
# 7
# 20
# 27, maybe AIC
# 28, F test based on tb

Ask Professor
# 20
# 26 from .5-.7 how will the cutoff change 


# 10

```{r}
mtcars
```

```{r}
lmod <- lm(mpg ~ hp, data = mtcars)
summary(lmod)
```

```{r}
# Make predictions using lmod
mtcars$pred <- predict(lmod, mtcars, type = "response")

# Make fake data to make the model line more smooth
mtcars_fake <- data.frame(hp = seq(50, 355, length.out = 1))
mtcars_fake$pred <- predict(lmod, mtcars_fake, type = "response")

# Overlay the model with the predictions
mtcars %>% 
  ggplot(aes(x = hp, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", method.args = list(family = "binomial")) +
  geom_line(data = mtcars_fake, aes(y = pred), linetype = "longdash")
```

# 11
```{r}
x <- matrix(c(8,4,0,-4,-8,7.8,9.0,10.2,11.0,11.7), ncol = 2, nrow = 5, byrow = F)

df <- as.data.frame(x)
```

```{r}
lmod <- lm(V2 ~ V1, data = df)
summary(lmod)
```

# 14
 chapter 2.2, pg 28


# 19

What is the lower 95% confidence interval (from the output above) for a 2 unit change in BNucl on the odds ratio scale? Use the normal approximation for the estimate – you can use 1.96 as the multiplier for a 95% confidence interval.

```{r}
BNucl_diff <- 2

library(faraway)
wbca

mod <- glm(Class ~ Adhes + BNucl, family = "binomial", data = wbca)
summary(mod)

exp_mod <- exp(coef(mod)); exp_mod
```

```{r}
tidy_mod <- broom::tidy(mod, conf.int = T) %>% 
  filter(term == "BNucl"); tidy_mod
```

```{r}
library(tidyverse)
tidy_mod %>% 
  select(estimate, conf.low, conf.high) %>%
  mutate_all(~exp(.*BNucl_diff))
```

```{r}
# check
odds_ratio <- exp(-0.7436686 * 2); odds_ratio

# Calcute log(OR) bc CI is calcuated in a linear scale
log_odds_ratio <- log(odds_ratio); log_odds_ratio

# Calcute exp(CI) to get CI in an odds ratio scale
CI_lower <- exp(log_odds_ratio - (1.96 * 0.07759735 * 2)); CI_lower
```